---
title: "Exploring Global Trends: Life Expectancy, GDP, and Population Analysis"
output: html_document
date: "2023-11-22"
---
The “gapminder.csv” file contains data about life expectancy, GDP per capita, and population by country. Use this dataset to create effective analysis/visualizations answering the following questions:

1. Investigate visually and comment on the distribution of world population as provided in this data
2. Establish the relationship between life expectancy and GDP per capita for different continents using a single suitable plot. Comment on the visual insights.
3. Extract the data corresponding to G7 (or G8) countries, then compare the life expectancy of the peoples' in these countries on one chart. Provide commentary interpretation on the obtained chart.

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
```

```{r}
# Read the dataset
fname = "gapminder.csv"
df = read.csv(fname)
class(df)
```
```{r}
# Return the first part of the dataframe
head(df)
```
```{r}
# Check the dimensions of the dataframe
dim(df)
```
```{r}
# Check the characteristics of the dataframe
glimpse(df)
```
```{r}
# Extract the summary statistics of the dataframe 
summary(df)
```
```{r}
# 1. Investigate visually and comment on the distribution of world population

# Create a histogram for world population
ggplot(df, aes(x = pop)) +
  geom_histogram(binwidth = 10000000, fill = "purple", color = "black", breaks = seq(0, 1500000000, by = 50000000)) +
  scale_x_continuous(labels = scales::comma) + 
  labs(x = "Population", y = "Frequency", title = "World Population Distribution")
```

# The above histogram shows that the population of most of the countries between the years 1952 - 2007 (a period of 55 years) is below 200,000,000. However, it may be noted that the population is distributed unevenly among countries. There are some countries with significantly larger populations, whereas, others have relatively smaller populations. The population sizes vary greatly, ranging from a few thousand (eg: Sao Tome and Djibouti) to billions (eg: China and India) in some countries. This uneven distribution creates a significant gap between the most and least populous countries.

```{r}
# 2. Establish the relationship between life expectancy and GDP per capita for different continents

# Create a scatter plot with life expectancy and GDP per capita
ggplot(df, aes(x = lifeExp, y = gdpPercap, color = continent)) + 
  geom_point() + 
  geom_smooth(col = 'black') + 
  labs(title = "Relationship between Life Expectancy and GDP per Capita by Continent", x = "Life Expectancy", y = "GDP per Capita")
```

# The scatter plot illustrates that there is a positive correlation between life expectancy and GDP per capita. Continents with higher life expectancy have a higher GDP per capita, especially in Europe and most of America where the life expectancy is between 70 to 85 years and GDP per capita is between 20,000 - 30,000. This may be due to the fact that such countries offer better access to healthcare, education, and other resources that contribute to longer life expectancies. However, there is a negative correlation in majority of Africa and some part of Asia where the life expectancy is mostly between 40 to 60 years and GDP per capita is less than 10,000. This is probably due to political instability, limited access to healthcare, economic disparities and environmental issues. It can also be seen that there are a few outliers in Asia where the GDP per capita exceeds 60,000, indicating that some countries perform exeptionally well as compared to others.

```{r}
# 3a. Extract the data corresponding to G7 countries 

# G7 countries list
g7_countries <- c("Canada", "France", "Germany", "Italy", "Japan", "United States", "United Kingdom")

# Filter the data for G7 countries
g7_data <- df %>% filter(country %in% g7_countries)
print(g7_data)
```

```{r}
# 3b. Compare the life expectancy of G7 countries in one chart

# Create a box plot for life expectancy of G7 countries
whisker_ends <- g7_data %>%
  group_by(country) %>%
  summarize(min = min(lifeExp), max = max(lifeExp))

ggplot(g7_data, aes(x = country, y = lifeExp)) +
  geom_boxplot(fill = "lightblue", color = "black", outlier.color = "red", outlier.shape = 16, outlier.size = 2, coef = 0) +
  geom_errorbar(data = whisker_ends, aes(ymin = min, ymax = max, x = country), width = 0.5, color = "black", linewidth = 1, inherit.aes = FALSE) +
  labs(title = "Life Expectancy Comparison of G7 Countries", x = "Country", y = "Life Expectancy")
```

# The above box plot depicts a notable trend in life expectancy. Across the G7 countries, there is a consistent increase in life expectancy from 1952 till 2007. This steady upward trajectory signifies substantial advancements in healthcare, living conditions, economic prosperity and societal improvements over time. However, there are a few differences among these countries such as Japan consistently maintains higher life expectancies as its median value is the highest (between 75 to 80 years) as compared to other G7 members. Addionally, the interquartile range of Japan is wider than the rest of the countries which suggests significant variation in life expectancy. This is mainly due to distinct healthcare systems or societal factors contributing to longer lifespans. It can also be seen that the median value of Germany and United Kingdom is the lowest (between 70 to 75 years) indicating low life expectancy in both of these countries. There are a few outliers above and below the median for each country, signifying exceptional cases that deviate from the general life expectancy trends.