---
title: "Statistical Analysis - Confidence Intervals and Hypothesis Testing"
author: "Sana Arif Raza"
date: "2024-10-14"
output:html_document: default
---

##### QUESTION #1: The ages in years file displays the ages of 50 randomly selected people in the civilian labor force. Determine a 95% confidence interval for the mean age, μ, of all people in the civilian labor force. Assume that the population standard deviation of the ages is 12.1 years.

```{r}
# Load the CSV file for Q1
ages_data = read.csv("Q1 Ages, in years.csv")
```

```{r}
# Access the 'AGES' column
ages = ages_data$AGES
```

```{r}
# Define the population standard deviation as given in question 1
pop_sd = 12.1
```

```{r}
# Calculate the sample mean
mean_age = mean(ages)
print(paste("Sample Mean Age:", mean_age))
```
```{r}
# Define the sample size
n = length(ages)
print(paste("Sample Size:", n))
```
```{r}
# Calculate the standard error of the mean
se = pop_sd / sqrt(n)
print(paste("Standard Error:", se))
```
```{r}
# The confidence interval is calculated using the z-score method since the population standard deviation is known

# Calculate the z-critical value for a 95% confidence interval
z_critical = qnorm(0.975)
print(paste("z-critical value:", z_critical))
```
```{r}
# Calculate the margin of error
margin_of_error = z_critical * se
print(paste("Margin of Error:", margin_of_error))
```
```{r}
# Calculate the 95% confidence interval and print the result
lower_bound = mean_age - margin_of_error
upper_bound = mean_age + margin_of_error
cat("95% Confidence Interval for the Mean Age: [", lower_bound, ",", upper_bound, "]\n")
```
##### Interpretation: The 95% confidence interval for the mean age means that we are 95% confident that the true population mean age of all people in the civilian labor force lies between 38.03 and 44.73 years.

##### QUESTION #2: The losses, in dollars, of 25 randomly selected pickpocket offenses are displayed in the losses for a sample of 25 pickpocket offenses file. Find a 95% confidence interval for the mean loss, μ, of all pickpocket offenses.

```{r}
# Load the CSV file for Q2
losses_data = read.csv("Q2 Losses ($) for a sample of 25 pickpocket offenses.csv")
```

```{r}
# Access the 'Losses' column
losses = losses_data$Losses
```

```{r}
# Calculate the sample mean
mean_loss = mean(losses)
print(paste("Sample Mean Loss:", mean_loss))
```
```{r}
# Calculate the sample standard deviation
sample_sd = sd(losses)
print(paste("Sample Standard Deviation:", sample_sd))
```
```{r}
# Define the sample size
n = length(losses)
print(paste("Sample Size:", n))
```
```{r}
# Calculate the standard error of the mean
se = sample_sd / sqrt(n)
print(paste("Standard Error:", se))
```
```{r}
# The confidence interval is calculated using the t-test method since the population standard deviation is not known

# Calculate the t-critical value for a 95% confidence interval (two-tailed)
t_critical = qt(0.975, df = n - 1)
print(paste("t-critical value:", t_critical))
```
```{r}
# Calculate the margin of error
margin_of_error = t_critical * se
print(paste("Margin of Error:", margin_of_error))
```
```{r}
# Calculate the 95% confidence interval and print the result
lower_bound = mean_loss - margin_of_error
upper_bound = mean_loss + margin_of_error
cat("95% Confidence Interval for the Mean Loss: [", lower_bound, ",", upper_bound, "]\n")
```
##### Interpretation: The 95% confidence interval for the mean loss means that we are 95% confident that the true average monetary loss per pickpocket offense falls between $405.08 and $621.56.

##### QUESTION #3: The table below shows the daily calcium intakes for a simple random sample of 18 adults with incomes below the poverty level. Decide, at the 5% significance level, whether the data provides sufficient evidence to conclude that the mean calcium intake of all adults with incomes below the poverty level is less than the RAI of 1000 mg per day. Assume that σ=188 mg.

```{r}
# Daily calcium intake data for 18 adults
calcium_intake = c(886, 633, 943, 847, 934, 841, 1193, 820, 774, 834, 
                    1050, 1058, 1192, 975, 1313, 872, 1079, 809)
```

```{r}
# Define the population standard deviation (σ) as given in question 3
sigma = 188
```

```{r}
# Calculate the sample size
n = length(calcium_intake)
print(paste("Sample Size:", n))
```
```{r}
# Calculate the sample mean
sample_mean = mean(calcium_intake)
print(paste("Sample Mean Calcium Intake:", sample_mean))
```
```{r}
# Define the hypothesized population mean (RAI) as given in question 3
mu_0 = 1000
```

```{r}
# Perform a one-sample z-test since the population standard deviation is known

# Calculate the z-test statistic
z = (sample_mean - mu_0) / (sigma / sqrt(n))
print(paste("Z value:", z))
```
```{r}
# Calculate the p-value for a one-tailed test
p_value = pnorm(z)
print(paste("P value:", p_value))
```
```{r}
# Define the significance level
alpha = 0.05

# Decision
if (p_value < alpha) {
  result = "Reject the null hypothesis: There is sufficient evidence to conclude that the mean calcium intake is less than 1000 mg."
} else {
  result = "Fail to reject the null hypothesis: There is not sufficient evidence to conclude that the mean calcium intake is less than 1000 mg."
}

# Print the result
cat(result, "\n")
```
##### Interpretation: The p-value (0.12) is greater than the significance level (0.05) which means that the result is not statistically significant at the 5% level. As a result, we fail to reject the null hypothesis. This means that based on the sample data, there is not sufficient evidence to conclude that the mean calcium intake of adults with incomes below the poverty level is less than 1000 mg per day.

##### QUESTION #4: The table below gives the pH levels of a sample of 15 lakes in the Southern Alps. Decide, at the 5% significance level, whether the data provides sufficient evidence to conclude that, on average, high mountain lakes in the Southern Alps are nonacidic.

```{r}
# pH levels for 15 lakes in the Southern Alps
pH_levels = c(7.2, 7.3, 6.1, 6.9, 6.6, 7.3, 6.3, 5.5, 6.3, 6.5, 
              5.7, 6.9, 6.7, 7.9, 5.8)

```

```{r}
# Calculate the sample mean
sample_mean = mean(pH_levels)
print(paste("Sample Mean of pH Level of Lakes:", sample_mean))
```
```{r}
# Hypotheses:
# Null hypothesis (H0): The average pH of the lakes is less than or equal to 7 (acidic)
# Alternative hypothesis (H1): The average pH of the lakes is greater than 7 (nonacidic)

# Perform a one-sample t-test with a hypothesized mean of 7
t_test = t.test(pH_levels, mu = 7, alternative = "less")
print(t_test)
```
```{r}
# Define the significance level
alpha = 0.05

# Decision
if (t_test$p.value < alpha) {
  result = "Reject the null hypothesis: There is sufficient evidence to conclude that the lakes are nonacidic."
} else {
  result = "Fail to reject the null hypothesis: There is not sufficient evidence to conclude that the lakes are nonacidic."
}

# Print the result
cat(result, "\n")
```
##### Interpretation: The p-value (0.02) is less than the significance level (0.05) which means that the result is statistically significant at the 5% level. As a result, we reject the null hypothesis. This means that based on the sample data, there is sufficient evidence to conclude that the mean pH level of high mountain lakes in the Southern Alps is less than 7. Therefore, the lakes tend to be acidic.
